---
title: "Assignment 1"
author: "Zack Wixom"
output: github_document
---

### WEEK 1 PROBLEMS

```{r}
# Load Packages
library(tidyverse)
library(rethinking)
```


1. _Suppose the globe tossing data (Chapter 2) had turned out to be 4 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book._

```
L W L W W L L W L L L L W L L 
```

```{r}
  # 1. Define the grid.
grid_appr <- tibble(grid = seq(from = 0, to = 1, length.out = 20)) %>% 
  # 2. Set the prior.
  mutate(prior = rep(1, 20)) %>% 
  # 3. Compute the likelihood.
  mutate(likelihood = dbinom(4, size = 15, prob = grid)) %>% 
  # 4. Compute the unstandardized posterior.
  mutate(unstd_posterior = likelihood * prior) %>% 
  # 5. Standardize the posterior.
  mutate(posterior = unstd_posterior / sum(unstd_posterior))

grid_appr

# Plot the grid approximation
ggplot(grid_appr, aes(x = grid, y = posterior)) +
  geom_line()

```

2. _Start over in 1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5. This corresponds to prior information that a majority of the Earth’s surface is water. What difference does the better prior make?_

```{r}

  # 1. Define the grid.
grid_appr <- tibble(grid = seq(from = 0, to = 1, length.out = 20)) %>% 
  # 2. Set the prior.
  mutate(prior = ifelse(grid < 0.5 , 0 , 1) ) %>% 
  # 3. Compute the likelihood.
  mutate(likelihood = dbinom(4, size = 15, prob = grid)) %>% 
  # 4. Compute the unstandardized posterior.
  mutate(unstd_posterior = likelihood * prior) %>% 
  # 5. Standardize the posterior.
  mutate(posterior = unstd_posterior / sum(unstd_posterior))

grid_appr

# Plot the grid approximation
ggplot(grid_appr, aes(x = grid, y = posterior)) +
  geom_line()

```


3. _For the posterior distribution from 2, compute 89% percentile and HPDI intervals. Compare the widths of these intervals._

```{r}

# Sample
grid_appr_sample <- sample(
  grid_appr$grid, 
  prob = grid_appr$posterior, 
  size = 10000, 
  replace = TRUE
)

# Percentile
pi_int <- PI(grid_appr_sample, prob = 0.89)

pi_int

# HPDI
hpdi_int <- HPDI(grid_appr_sample, prob = 0.89)

hpdi_int


```

_Which is wider? Why? _

- The Percentile Interval is wider and I think that is because the HPDI takes a denser interval by including the highest point.

_If you had only the information in the interval, what might you misunderstand about the shape of the posterior distribution?_

Since both the PI and HPDI are pretty similar that that means that the shape is more or less bell-shaped (or at least that's what it sounded like happens in the book), however, I am sure that the shape can be misunderstood by just looking at these numbers, you would need to plot them to figure out the true shape.

_Questions from the assignment:_

- How do you determine the size of the grid. In the book he did 20 with only 9 tosses. Is it normal to make your grid larger than your number of observed data? Why not make them the same?

- Does the size matter when sampling form the posterior?

- How do you plot the PI and HPDI on a existing graph? Is there a way to add the shading?

- I am slightly confused about how the PI and HPDI can help you determine the shape of the distrubtion. Maybe there is something I missed from the reading.







