---
title: "Connecting to Fortnite API"
author: "Zack"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/"
)
```

I am going to try to connect to an API for [Fortnite Tracker](https://fortnitetracker.com/). This API gives you stats on every player you have the gamertag of. It updates frequently with current matches played.

Hopefully I can connect to this API becuase it seems like the best option for getting the most relevant data I want to analyze. 

However, I couldn't find any documentation on connecting to it with R. So I am going to use similar code for a different API that uses `httr` package to connect to an API and see if I can connect to Fortnite Tracker that way.

## Attempt to Connect

the `httr` package has documentation to connect to many different API's from website called RapidAPI. I am using this [example](https://rapidapi.com/elreco/api/fortnite-api?endpoint=apiendpoint_a787830a-94d7-4d1b-a625-031462c97605) of a different Forntie API call.

```{r httr-example, include = FALSE}
# Packages
library(tidyverse)
library(httr)
# 
# url <- "https://fortnite-api.p.rapidapi.com/stats/zRotation"
# 
# response <- VERB("GET", url, add_headers(x_rapidapi_key = '751fdcf83amshb438c02e00cb7f9p189025jsne07cf0e465d3', x_rapidapi_host = 'fortnite-api.p.rapidapi.com'), content_type("application/octet-stream"))
# 
# content(response, "text")

```

After some messing around I wasn't able to get it to work so I found another resource [here](https://www.dataquest.io/blog/r-api-tutorial/). This says I can use a package called `jsonlite` along with httr to access API's. So I will try that out, for simplicity I am just going to get data from the in game store. 

```{r setup, eval = FALSE}
# Packages
library(tidyverse)
library(httr)
library(jsonlite)

# Url of API
url <- "https://api.fortnitetracker.com/v1/store"

# # Save API Key to Private Folder
# saveRDS(api_key, here::here("Private", "api_key.rds"))

# Read in API Key
api_key <- readRDS(here::here("Private", "api_key.rds"))

```

Now that I have my packages loaded, key saved, and url added I can run the `httr` code.

```{r httr practice}

# Connect to API
resp <- httr::GET(url,
    add_headers('TRN-API-Key' = api_key))

# Check for any errors
http_error(resp) 

# Get content use "parsed" to make it an r object
result <- httr::content(resp, as = "parsed")

# Convert to dataset
data <- data.frame(result)

data

```

It worked!! So now let me test getting some player stats. Eventually I will need to make a list of player names that I can look through and create a dataset of data for players.

```{r player stat test}

# base Url of API
url <- "https://api.fortnitetracker.com/v1/profile"

# Platform
platform <- "gamepad"

# GamerTag
gamertag <- "princesstress"

# Connect to API
resp <- httr::GET(url = paste(url, platform, gamertag, sep = "/"),
    add_headers('TRN-API-Key' = api_key))

# Check for any errors
http_error(resp) 

# str(resp)

# Get content use "parsed" to make it an r object
result <- httr::content(resp, as = "parse")

# Convert to dataset
data <- data.frame(result)


```

The result is pretty complex but has all the information we need. We just need to clean it up and make it easier to understand. first I am going to just get the lifetime stats of the player along with their identifying information.

There are levels and groups to the data. 

**Overview Level** --> identifying information
- account ID
- Avatar
- Platform ID
- Platform Name
- Platform Name Long
- Epic User Handle

**Stats Level** --> Recent Game Stats

Stats for every time of game mode:
- p2    = Solo Matches
- p10   = Duos Matches
- p9    = Squad Matches
- ltm   = Limited Time Matches (special game modes, etc)

Within each game mode there are these statistics:
- trnRating     = Tracker Network Rating
- score         = Score
- top1          = Wins (Victory Royales)
- top3          = Top 3 placement
- top5          = Top 5 
- top6          = Top 6 
- top10         = Top 10 
- top12         = Top 12
- top25         = Top 25
- kd            = Kill/Death Ratio
- winRatio      = Win Ratio
- matches       = Number of Matches
- kills         = Number of Eliminations
- minutesPlayed = Time Played in Minutes
- kpm           = Kills per Minute
- kpg           = Kills per Match
- avgTimePlayed = Average Match Time
- scorePerMatch = Score per Match
- scorePerMin   = Score per Minute

**Lifetime Stats** --> Lifetime Overall Stats

The first six levels are all about placement in matches:
- 1     = Top 5 placement
- 2     = Top 3
- 3     = Top 6
- 4     = Top 10
- 5     = Top 12
- 6     = Top 25

The next six are various other stats:
- 7     = Score (of points from in game XP)
- 8     = Matches Played
- 9     = Wins (Victory Royales)
- 10    = Win Ratio %
- 11    = Eliminations
- 12    = Kills/Death Ratio %

**Recent Matches** --> Recent Match Session Stats

### Extracting Season Solo Match Stats

My plan is to extract each of these datasets so that I can tidy up the data.

```{r id data}
# Create data with just Idenfitication Data
id_data <- data.frame(result[1:6]) 

# username
id <- result$epicUserHandle

```

```{r solo stats test}

# Loop to get all the stats for Solo
for (i in seq(19)) {
  x <- result$stats$p2[i]
  
  y <- x[[1]]
  
  z <- y[[4]]
  
  a <- data.frame(z)
  
  names(a)[names(a) == "z"] <- y[[2]]
  
  if(i == 1){
    solo_data <- a %>% 
      mutate(
        id = id
      )
  }else{
    solo_data <- cbind(solo_data, a)
  }
  solo_data
}

solo_data
```

Perfect I was able to create a loop that allowed me to grab all of the stats for the Solo gamemode. 

Now let me do the same thing but for the lifetime stats.

```{r lifetime stat test}
# username
id <- result$epicUserHandle

# Loop to get all the stats for Solo
for (i in seq(12)) {
  stat_level <- result$lifeTimeStats[[i]]
  
  stat_label <- stat_level[1]
  
  stat_value <- stat_level[2]
  
  stat_df <- data.frame(stat_value)
  
  names(stat_df)[names(stat_df) == "value"] <- stat_label
  
  if(i == 1){
    life_stat <- stat_df %>% 
      mutate(
        id = id
      )
  }else{
    life_stat <- cbind(life_stat, stat_df)
  }
  life_stat
}

life_stat
```

Now I have the lifetime stats as well using a loop. 

I realized that I need a list of gamertags that I will use to grab stats for. So I am going to add that dataset which was found on [Kaggle](https://www.kaggle.com/jackdaoud/esports-earnings-for-players-teams-by-game?select=highest_earning_players.csv)

This dataset includes total earnings of top 100 players of various games. I will filter by Fortnite and then use those names in my data pull.

*Note: data has been updated last 2 months*

```{r earnings data}

# Pull earnings data
earnings_data <- read_csv(here::here("Project", "Data", "highest_earning_players.csv")) %>% 
  filter(Game == "Fortnite") %>% 
  select(CurrentHandle, TotalUSDPrize)

# Get list of player names
player_names <- earnings_data$CurrentHandle

# test to see if I can index
player_names[1]
```

Now I have a data set that has the current gamertag of the player and their total earnings in USD.

```{r extract data}
# Platform
platform <- "kbm"

# GamerTag
gamertag <- "bugha"

# Connect to API
resp <- httr::GET(url = paste(url, platform, gamertag, sep = "/"),
    add_headers('TRN-API-Key' = api_key))

# Check for any errors
http_error(resp) 

# str(resp)

# Get content use "parsed" to make it an r object
result <- httr::content(resp, as = "parse")

# username
id <- result$epicUserHandle

# Loop to get all the stats for Solo
for (i in seq(12)) {
  stat_level <- result$lifeTimeStats[[i]]
  
  stat_label <- stat_level[1]
  
  stat_value <- stat_level[2]
  
  stat_df <- data.frame(stat_value)
  
  names(stat_df)[names(stat_df) == "value"] <- stat_label
  
  if(i == 1){
    life_stat <- stat_df %>% 
      mutate(
        id = id
      )
  }else{
    life_stat <- cbind(life_stat, stat_df)
  }
  life_stat
}

life_stat

```

So I have run into the issue that not all fortnite players have the stats on the API. It has some but it isn't completely consistent. So I might need to change up my data source.


