---
title: "Twitch Streamer Analysis"
author: "Zack Wixom"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/Project/"
)
```


I am going to try and look at just broad Twitch streamer data to build a model off of it. I have identified a dataset from Kaggle that gives me lots of variables that I could help me model on minutes streamed. 

There is also a Twitch API that might be better cause I would like more variables to look at such as the main game they stream, etc. But for now I am looking at just the stats.

```{r}
# Packages
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(rethinking)
library(dagitty)

# Load Data
twitch <- read_csv(here::here("Project/Data", "twitchdata-update.csv"))

twitch

```

## The Outcome Variable

There are two variables that would be interesting to model on: `Watch time(Minutes)` is the overall time that viewers have watched their videos and `Followers` the number of followers on Twitch. `Followers` is interesting because it is something that might be a bigger overall outcome that Twitch streamers are looking for because followers turn into subs which turn into revenue.

## DAG

I want to understand what variables I want to include and condition on so I will build a DAG. I am just going to use the continuous variables with `Partnered` variable instead of using all the other discrete variables.

```{r dag1}
# Make Dag to look at what to condition on
dag_01 <- dagitty( "dag {
  A -> F
  W -> F
  W -> A
  S -> F
  S -> W
  S -> A
  Pa -> S
  Pa -> W
}")

plot(dag_01)


```

There is a lot going on here with all these variables. The main paths I want to explore are A -> F, W -> F and S -> F. In order to do this I will need to condition on Pa for the mode, since it is a back door.

## Running Models

First I want to check out all the variables and check out if I should log any of them or make any index variables.

```{r prep data}
# Standardize  and Index Variables
twitch1 <- tibble(
    followers = standardize(twitch$Followers),
    average_viewers = standardize(twitch$`Average viewers`),
    peak = standardize(twitch$`Peak viewers`),
    watchtime = standardize(twitch$`Watch time(Minutes)`),
    partnered = as.integer(twitch$Partnered),
    mature = as.integer(twitch$Mature),
    language = as.integer(factor(twitch$Language))
)

dens(twitch1$followers)
dens(twitch1$average_viewers)
dens(twitch1$peak)
dens(twitch1$partnered)
dens(twitch1$mature)
dens(twitch1$language)

```

All the continuous variables are skewed so I will need to log them to get an accurate distribution. However I also noticed that almost all of the channels are Partnered and so that means that there isn't going to be any real effect happening so I need to update my DAG without Pa.

```{r dag2}
# Make Dag to look at what to condition on
dag_02 <- dagitty( "dag {
  A -> F
  W -> F
  W -> A
  S -> F
  S -> W
  S -> A
}")

plot(dag_02)


```

```{r prep data contd}

# Standardize  and Index Variables
twitch2 <- tibble(
    followers = standardize(log(twitch$Followers)),
    average_viewers = standardize(log(twitch$`Average viewers`)),
    peak = standardize(log(twitch$`Peak viewers`)),
    watchtime = standardize(log(twitch$`Watch time(Minutes)`))
)

# Check distributions
dens(twitch2$followers)
dens(twitch2$average_viewers)
dens(twitch2$peak)
```

Now I can create my first model

```{r model1}
# define average Average Viewers
xbar <- mean(twitch2$average_viewers)

# Fit Model
m1 <- quap(
  alist(
    followers ~ dnorm(mu, sigma),
    mu <- a + b_avgviewers*average_viewers + b_peak * peak + b_watchtime*watchtime,
    a ~ dnorm(0, 1),
    b_avgviewers ~ dlnorm(0, .5),
    b_peak ~ dlnorm(0, .5),
    b_watchtime ~ dlnorm(0, .5),
    sigma ~ dexp(1)
  ), data = twitch2
)

plot(precis(m1))

```

So it looks like that these three variables are very impacting for increasing followers on Twitch. This isn't much of a surprise but it will be interesting to add in more variables such as Mature and Language

```{r dag3}
# Make Dag to look at what to condition on
dag_03 <- dagitty( "dag {
  A -> F
  W -> F
  W -> A
  S -> F
  S -> W
  S -> A
  M -> W
  L -> W
}")

plot(dag_03)


```

Well I am not sure how to deal with what variables are influencing each other so I am going to assume just that `Language` and `Mature` discrete variables are influencing `Watch time (minutes)` becuase the older the audience probably the more they watch so `Mature` might indicate age (on average) also I think that `Language` could influence `Watch time (minutes)`because there might be more streamers in certain languages than others so certain languages like English might have more watchers and increase the overall `Watch time (minutes)`.

```{r data prep 2}
# Standardize  and Index Variables
twitch3 <- tibble(
    followers = standardize(log(twitch$Followers)),
    average_viewers = standardize(log(twitch$`Average viewers`)),
    peak = standardize(log(twitch$`Peak viewers`)),
    watchtime = standardize(log(twitch$`Watch time(Minutes)`)),
    mature = as.integer(twitch$Mature),
    language = as.integer(factor(twitch$Language))
)

# Check distributions
dens(twitch3$mature)
dens(twitch3$language)

```

Checking out the Density os the `mature` and `language` variables for this model we see that most streamers are not mature and speak English. So we will see if this influences the number of followers they have.

```{r model2}
# Fit Model
m2 <- quap(
  alist(
    followers ~ dnorm(mu, sigma),
    mu <- a[language] + b_avgviewers*average_viewers + b_peak * peak + b_watchtime*watchtime,
    a[language] ~ dnorm(0, 1),
    b_avgviewers ~ dlnorm(0, .5),
    b_peak ~ dlnorm(0, .5),
    b_watchtime ~ dlnorm(0, .5),
    sigma ~ dexp(1)
  ), data = twitch3
)

plot(precis(m2, depth = 2))

```

So there are definitely some languages that indicate more followers or less. This seems fairly obvious since the languages that are less common in general there will be less big streamers and so they will have less followers.

Let's do the same thing but with `mature` variables.

```{r model3}
# Fit Model
m3 <- quap(
  alist(
    followers ~ dnorm(mu, sigma),
    mu <- a[mature] + b_avgviewers*average_viewers + b_peak * peak + b_watchtime*watchtime,
    a[mature] ~ dnorm(0, 1),
    b_avgviewers ~ dlnorm(0, .5),
    b_peak ~ dlnorm(0, .5),
    b_watchtime ~ dlnorm(0, .5),
    sigma ~ dexp(1)
  ), data = twitch3
)

plot(precis(m3, depth = 2))

```

Looks like `mature` is not giving us any significant changes with followers.

## Prior Predicitve Checks

It will be useful to see if I am using good Priors right now before moving onto `ulam()` function for my models.



