---
title: "Twitch Streamer Analysis"
author: "Zack Wixom"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/"
)
```


I am going to try and look at just broad Twitch streamer data to build a model off of it. I have identified a dataset from Kaggle that gives me lots of variables that I could help me model on minutes streamed. 

There is also a Twitch API that might be better cause I would like more variables to look at such as the main game they stream, etc. But for now I am looking at just the stats.

```{r}
# Packages
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(rethinking)

# Load Data
twitch <- read_csv(here::here("Project/Data", "twitchdata-update.csv"))

twitch

```

## The Outcome Variable

There are two variables that would be interesting to model on: `Watch time(Minutes)` is the overall time that viewers have watched their videos and `Followers` the number of followers on Twitch. `Followers` is interesting because it is something that might be a bigger overall outcome that Twitch streamers are looking for becuase followers turn into subs which turn into revenue.

## Running a Linear Model - Milestone 

For a major milestone of the project, I need to build a linear model. This was found in our week 3 reading and assignment 2. 

So I am going to build a simple model on `Followers` with the explanatory variable of `Average Viewers`. 

```{r}
# Standardize  and Index Variables
twitch1 <- list(
    followers = standardize(twitch$Followers),
    average_viewers = standardize(twitch$`Average viewers`)
)

dens(twitch1$followers)

dens(twitch1$average_viewers)

```


```{r}
# define average Average Viewers
xbar <- mean(twitch1$average_viewers)

# Fit Model
m1 <- quap(
  alist(
    followers ~ dnorm(mu, sigma),
    mu <- a + b*(average_viewers - xbar),
    a ~ dnorm(0, 1),
    b ~ dnorm(0, 0.2),
    sigma ~ dexp(1)
  ), data = twitch1
)

plot(precis(m1))

```








