---
title: "Twitch Streamer Analysis"
author: "Zack Wixom"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/Project/"
)
```


I am going to try and look at just broad Twitch streamer data to build a model off of it. I have identified a dataset from Kaggle that gives me lots of variables that I could help me model on minutes streamed. 

There is also a Twitch API that might be better cause I would like more variables to look at such as the main game they stream, etc. But for now I am looking at just the stats.

```{r}
# Packages
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(rethinking)
library(dagitty)

# Load Data
twitch <- read_csv(here::here("Project/Data", "twitchdata-update.csv"))

twitch

```

## The Outcome Variable

There are two variables that would be interesting to model on: `Watch time(Minutes)` is the overall time that viewers have watched their videos and `Followers` the number of followers on Twitch. `Followers` is interesting because it is something that might be a bigger overall outcome that Twitch streamers are looking for because followers turn into subs which turn into revenue.

## DAG

I want to understand what variables I want to include and condition on so I will build a DAG.

```{r dag}
# Standardize Variables
d <- twitch %>% 
  mutate(
    F = standardize(Followers),
    A = standardize(`Average viewers`),
    W = standardize(`Watch time(Minutes)`),
    S = standardize(`Stream time(minutes)`),
    P = standardize(`Peak viewers`),
    Pa = as.integer(Partnered)
  )


# Make Dag to look at what to condition on
dag_01 <- dagitty( "dag {
  A -> F
  W -> F
  W -> A
  S -> F
  S -> W
  S -> A
  Pa -> S
  Pa -> W
}")

plot(dag_01)


```


## Running a Linear Model - Milestone 

For a major milestone of the project, I need to build a linear model. This was found in our week 3 reading and assignment 2. 

So I am going to build a simple model on `Followers` with the explanatory variable of `Average Viewers`. 

```{r}
# Standardize  and Index Variables
twitch1 <- list(
    followers = standardize(twitch$Followers),
    average_viewers = standardize(twitch$`Average viewers`),
    peak = standardize(twitch$`Peak viewers`),
    watchtime = standardize(twitch$`Watch time(Minutes)`)
)

dens(twitch1$followers)

dens(twitch1$average_viewers)

dens(twitch1$peak)

```


```{r}
# define average Average Viewers
xbar <- mean(twitch1$followers)

# Fit Model
m1 <- quap(
  alist(
    average_viewers ~ dnorm(mu, sigma),
    mu <- beta0 + beta1*(average_viewers - xbar),
    beta0 ~ dnorm(0, 1),
    beta1 ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  ), data = twitch1
)

plot(precis(m1))

```

```{r}
# 1. Specify the new data we'd like to predict new heights.
new_average_viewers <- seq(from = 0, to = 1)

# 2. Simulate predicted heights using the likelihood height_i ∼ Normal(mu_i, sigma).
posterior_pd <- sim(m4_2, data = list(weight = new_weights))

# Plot the posterior prediction distribution and compare with the data.
hist(posterior_pd)
hist(data$height)
```



```{r}
diag(vcov(m1))
cov2cor(vcov(m1))
```
```{r}
# Sample from the posterior.
posterior <- extract.samples(m1, n = 10000)
# Some possible summaries.
precis(posterior)
as_tibble(posterior) %>% 
  pivot_longer(
    a:sigma, 
    names_to = "parameters", 
    values_to = "draws"
  ) %>% 
  ggplot(aes(x = draws)) +
  geom_histogram() +
  facet_wrap(~ parameters, scales = "free")

```



```{r}
# 1. Specify the number of lines to simulate.
N <- 100
prior_pd <- tibble(
  # 2. Specify the line number.
  n = 1:N,
  # 3. Simulate values of beta0 and beta1 from their priors.
  beta0 = rnorm(N, 0, 1),
  beta1 = rlnorm(N, 0, .5)   
) %>%
  # 4. Create two rows for each line, one for each end of the range.
  expand(nesting(n, beta0, beta1), average_viewers = range(twitch1$average_viewers)) %>% 
  mutate(
    # 5. Simulate average height using the linear model:
    followers = beta0 + beta1 * (average_viewers - mean(twitch1$average_viewers)),
  )
# Plot the prior predictive distribution regression lines.
prior_pd %>% 
  ggplot(aes(x = average_viewers, y = followers, group = n)) +
  geom_line(alpha = 0.10)
```




