---
title: "Smash Data Analysis"
subtitle: "Collecting and Cleaning Data"
author: "Zack Wixom"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/Project/Smash"
)
```

```{r packages, eval = FALSE, message = FALSE}
# Load Packages
library(tidyverse)
library(tidytext)
library(textdata)
library(rvest)

```

So my first attempt to get data on Fortnite players to understand earnings and victory royales didn't work. My API didn't have sufficient data for the players I wanted to analyze. So I am going to try to go a different approach with Super Smash Bros.

I think that I can build a model to understand the relationship between character stats and their win rates. I want to see if frame data, weight, speed, etc, will increase or decrease chance of winning a match against another character. Ideally I will be able to pit two characters up against each other and see which one wins.

## Collecting Data

### Frame Data

In order to do this I need two sources of data: Character data and Match data. I know of a [website](https://ultimateframedata.com/) that has frame data for every character and I know of at least one data set of match results data. I am not sure if I can get more but the more the better so I can see the match ups for every character so I can make some kind of comparison between the characters. Not sure if this is possible but I will see. 

```{r webscrapping frame}

## Web scrapping for Frame Data
tables <- read_html("https://ultimateframedata.com/stats.php") %>%
  html_nodes("table")

# # Practice table webscrape
# air_acceleration <- tables %>% 
#   .[1] %>%
#   html_table(fill = TRUE)

# Loop to grab all data tables
for (i in 1:14) {
  x <- tables %>% 
  .[i] %>%
  html_table(fill = TRUE)
  
  write_csv(as.data.frame(x), paste("smash", i, ".csv", sep = ""))
}

# Explore the varialbe names
acceleration <- read.csv(here::here("Project", "Data", "smash1_acceleration.csv"))
airspeed <- read.csv(here::here("Project", "Data", "smash2_airspeed.csv"))
head(acceleration)
head(airspeed)


```

Every Data table has a identifier, `Character`, variable. This will be used to join together the datasets. However, there are some variable names that are repeated but need to be fixed before joining together. For example, every dataset comes with the ranking of the characters for that statistic. So that will need to be changed so that we can differentiate before each table's ranking.

I also need to clean up the `Character` variable a bit. Some of the tables have echo fighters together and some have them separate. The final dataset will have them separate since people choose one echo fighter over the other for various reasons (i.e. Dark Samus over Samus, Daisy over Peach, etc.).

```{r load data}

# Load in Air Acceleration Data >>> Needed to define variables better to avoid confusion
acceleration <- read.csv(here::here("Project", "Data", "smash1_acceleration.csv")) %>%
  rename(
    rank_accel = Rank,
    base_accel = Base,
    additional_accel = Additional,
    max_accel = Max
  )

# acceleration <- read.csv(here::here("Project", "Data", "smash1_acceleration.csv"))

# Fix Names
acceleration$Character[acceleration$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
acceleration$Character[acceleration$Character == "DK"] <- "Donkey Kong"
acceleration$Character[acceleration$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
acceleration$Character[acceleration$Character == "DK"] <- "Donkey Kong"
acceleration$Character[acceleration$Character == "P.T. [Charizard]"] <- "Charizard"
acceleration$Character[acceleration$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
acceleration$Character[acceleration$Character == "P.T. [Squirtle]"] <- "Squirtle"
acceleration$Character[acceleration$Character == "PAC-MAN"] <- "Pac-Man"
acceleration$Character[acceleration$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
acceleration$Character[acceleration$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
acceleration$Character[acceleration$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
acceleration$Character[acceleration$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
acceleration$Character[acceleration$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
acceleration$Character[acceleration$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
acceleration$Character[acceleration$Character == "Dedede"] <- "King Dedede"
acceleration$Character[acceleration$Character == "Metaknight"] <- "Meta Knight"
acceleration$Character[acceleration$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
acceleration <- acceleration %>% 
  separate_rows(Character, sep = "/")

# Write Data Locally
write_csv(acceleration, here::here("Project", "Data", "smash1_acceleration.csv"))

# Load in Air Speed Data
airspeed <- read.csv(here::here("Project", "Data", "smash2_airspeed.csv")) %>% 
  rename(
    rank_airspeed = Rank
  )

# airspeed <- read.csv(here::here("Project", "Data", "smash2_airspeed.csv"))

# Fix Names
airspeed$Character[airspeed$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
airspeed$Character[airspeed$Character == "DK"] <- "Donkey Kong"
airspeed$Character[airspeed$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
airspeed$Character[airspeed$Character == "DK"] <- "Donkey Kong"
airspeed$Character[airspeed$Character == "P.T. [Charizard]"] <- "Charizard"
airspeed$Character[airspeed$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
airspeed$Character[airspeed$Character == "P.T. [Squirtle]"] <- "Squirtle"
airspeed$Character[airspeed$Character == "PAC-MAN"] <- "Pac-Man"
airspeed$Character[airspeed$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
airspeed$Character[airspeed$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
airspeed$Character[airspeed$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
airspeed$Character[airspeed$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
airspeed$Character[airspeed$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
airspeed$Character[airspeed$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
airspeed$Character[airspeed$Character == "Dedede"] <- "King Dedede"
airspeed$Character[airspeed$Character == "Metaknight"] <- "Meta Knight"
airspeed$Character[airspeed$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
airspeed <- airspeed %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(airspeed, here::here("Project", "Data", "smash2_airspeed.csv"))

# Load in Fall Speed Data
fallspeed <- read.csv(here::here("Project", "Data", "smash3_fallspeed.csv")) %>% 
  rename(
    rank_fallspeed = Rank
  )

# fallspeed <- read.csv(here::here("Project", "Data", "smash3_fallspeed.csv"))

# Fix Names
fallspeed$Character[fallspeed$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
fallspeed$Character[fallspeed$Character == "DK"] <- "Donkey Kong"
fallspeed$Character[fallspeed$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
fallspeed$Character[fallspeed$Character == "DK"] <- "Donkey Kong"
fallspeed$Character[fallspeed$Character == "P.T. [Charizard]"] <- "Charizard"
fallspeed$Character[fallspeed$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
fallspeed$Character[fallspeed$Character == "P.T. [Squirtle]"] <- "Squirtle"
fallspeed$Character[fallspeed$Character == "PAC-MAN"] <- "Pac-Man"
fallspeed$Character[fallspeed$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
fallspeed$Character[fallspeed$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
fallspeed$Character[fallspeed$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
fallspeed$Character[fallspeed$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
fallspeed$Character[fallspeed$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
fallspeed$Character[fallspeed$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
fallspeed$Character[fallspeed$Character == "Dedede"] <- "King Dedede"
fallspeed$Character[fallspeed$Character == "Metaknight"] <- "Meta Knight"
fallspeed$Character[fallspeed$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
fallspeed <- fallspeed %>% 
  separate_rows(Character, sep = "/")

# Write Data Locally
write_csv(fallspeed, here::here("Project", "Data", "smash3_fallspeed.csv"))

# Load in Gravity Data
gravity <- read.csv(here::here("Project", "Data", "smash4_gravity.csv"))

# Fix Names
gravity$Character[gravity$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
gravity$Character[gravity$Character == "DK"] <- "Donkey Kong"
gravity$Character[gravity$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
gravity$Character[gravity$Character == "DK"] <- "Donkey Kong"
gravity$Character[gravity$Character == "P.T. [Charizard]"] <- "Charizard"
gravity$Character[gravity$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
gravity$Character[gravity$Character == "P.T. [Squirtle]"] <- "Squirtle"
gravity$Character[gravity$Character == "PAC-MAN"] <- "Pac-Man"
gravity$Character[gravity$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
gravity$Character[gravity$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
gravity$Character[gravity$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
gravity$Character[gravity$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
gravity$Character[gravity$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
gravity$Character[gravity$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
gravity$Character[gravity$Character == "Dedede"] <- "King Dedede"
gravity$Character[gravity$Character == "Metaknight"] <- "Meta Knight"
gravity$Character[gravity$Character =="Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
gravity <- gravity %>% 
  separate_rows(Character, sep = "/")

# Write Data Locally
write_csv(gravity, here::here("Project", "Data", "smash4_gravity.csv"))

# Load in Jump Height Data
jumpheight <- read.csv(here::here("Project", "Data", "smash5_jumpheight.csv")) %>% 
  rename(
    shorthop_height = Short.Hop,
    fullhop_height = Full.Hop,
    airjump_height = Air.Jump
  )

# jumpheight <- read.csv(here::here("Project", "Data", "smash5_jumpheight.csv"))

# Fix Names
jumpheight$Character[jumpheight$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
jumpheight$Character[jumpheight$Character == "DK"] <- "Donkey Kong"
jumpheight$Character[jumpheight$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
jumpheight$Character[jumpheight$Character == "DK"] <- "Donkey Kong"
jumpheight$Character[jumpheight$Character == "P.T. [Charizard]"] <- "Charizard"
jumpheight$Character[jumpheight$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
jumpheight$Character[jumpheight$Character == "P.T. [Squirtle]"] <- "Squirtle"
jumpheight$Character[jumpheight$Character == "PAC-MAN"] <- "Pac-Man"
jumpheight$Character[jumpheight$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
jumpheight$Character[jumpheight$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
jumpheight$Character[jumpheight$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
jumpheight$Character[jumpheight$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
jumpheight$Character[jumpheight$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
jumpheight$Character[jumpheight$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
jumpheight$Character[jumpheight$Character == "Dedede"] <- "King Dedede"
jumpheight$Character[jumpheight$Character == "Metaknight"] <- "Meta Knight"
jumpheight$Character[jumpheight$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
jumpheight <- jumpheight %>% 
  separate_rows(Character, sep = "/")

# Write Data Locally
write_csv(jumpheight, here::here("Project", "Data", "smash5_jumpheight.csv"))

# Load in Jump Duration Data
jumpduration <- read.csv(here::here("Project", "Data", "smash6_jumpduration.csv")) %>% 
  rename(
    shorthop_dur = Short.Hop,
    fullhop_dur = Full.Hop
  )

# jumpduration <- read.csv(here::here("Project", "Data", "smash6_jumpduration.csv"))

# Fix Names
jumpduration$Character[jumpduration$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
jumpduration$Character[jumpduration$Character == "DK"] <- "Donkey Kong"
jumpduration$Character[jumpduration$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
jumpduration$Character[jumpduration$Character == "DK"] <- "Donkey Kong"
jumpduration$Character[jumpduration$Character == "P.T. [Charizard]"] <- "Charizard"
jumpduration$Character[jumpduration$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
jumpduration$Character[jumpduration$Character == "P.T. [Squirtle]"] <- "Squirtle"
jumpduration$Character[jumpduration$Character == "PAC-MAN"] <- "Pac-Man"
jumpduration$Character[jumpduration$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
jumpduration$Character[jumpduration$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
jumpduration$Character[jumpduration$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
jumpduration$Character[jumpduration$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
jumpduration$Character[jumpduration$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
jumpduration$Character[jumpduration$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
jumpduration$Character[jumpduration$Character == "Dedede"] <- "King Dedede"
jumpduration$Character[jumpduration$Character == "Metaknight"] <- "Meta Knight"
jumpduration$Character[jumpduration$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
jumpduration <- jumpduration %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(jumpduration, here::here("Project", "Data", "smash6_jumpduration.csv"))

# Load in Weight Data
weight <- read.csv(here::here("Project", "Data", "smash7_weight.csv")) %>% 
  rename(
    rank_weight = Rank
  )

weight <- read.csv(here::here("Project", "Data", "smash7_weight.csv"))

# Fix Names
weight$Character[weight$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
weight$Character[weight$Character == "DK"] <- "Donkey Kong"
weight$Character[weight$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
weight$Character[weight$Character == "DK"] <- "Donkey Kong"
weight$Character[weight$Character == "P.T. [Charizard]"] <- "Charizard"
weight$Character[weight$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
weight$Character[weight$Character == "P.T. [Squirtle]"] <- "Squirtle"
weight$Character[weight$Character == "PAC-MAN"] <- "Pac-Man"
weight$Character[weight$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
weight$Character[weight$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
weight$Character[weight$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
weight$Character[weight$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
weight$Character[weight$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
weight$Character[weight$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
weight$Character[weight$Character == "Dedede"] <- "King Dedede"
weight$Character[weight$Character == "Metaknight"] <- "Meta Knight"
weight$Character[weight$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
weight <- weight %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(weight, here::here("Project", "Data", "smash7_weight.csv"))

# Load in Landing Data
landing <- read.csv(here::here("Project", "Data", "smash8_landing.csv")) %>% 
  separate_rows(Character, sep = "/")

# landing <- read.csv(here::here("Project", "Data", "smash8_landing.csv"))

# Fix Names
landing$Character[landing$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
landing$Character[landing$Character == "DK"] <- "Donkey Kong"
landing$Character[landing$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
landing$Character[landing$Character == "DK"] <- "Donkey Kong"
landing$Character[landing$Character == "P.T. [Charizard]"] <- "Charizard"
landing$Character[landing$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
landing$Character[landing$Character == "P.T. [Squirtle]"] <- "Squirtle"
landing$Character[landing$Character == "PAC-MAN"] <- "Pac-Man"
landing$Character[landing$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
landing$Character[landing$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
landing$Character[landing$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
landing$Character[landing$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
landing$Character[landing$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
landing$Character[landing$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
landing$Character[landing$Character == "Dedede"] <- "King Dedede"
landing$Character[landing$Character == "Metaknight"] <- "Meta Knight"
landing$Character[landing$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
landing <- landing %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(landing, here::here("Project", "Data", "smash8_weight.csv"))

# Load in Walk Speed Data
walkspeed <- read.csv(here::here("Project", "Data", "smash9_walkspeed.csv")) %>% 
  rename(
    rank_walkspeed = Rank
  )

# walkspeed <- read.csv(here::here("Project", "Data", "smash9_walkspeed.csv"))

# Fix Names
walkspeed$Character[walkspeed$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
walkspeed$Character[walkspeed$Character == "DK"] <- "Donkey Kong"
walkspeed$Character[walkspeed$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
walkspeed$Character[walkspeed$Character == "DK"] <- "Donkey Kong"
walkspeed$Character[walkspeed$Character == "P.T. [Charizard]"] <- "Charizard"
walkspeed$Character[walkspeed$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
walkspeed$Character[walkspeed$Character == "P.T. [Squirtle]"] <- "Squirtle"
walkspeed$Character[walkspeed$Character == "PAC-MAN"] <- "Pac-Man"
walkspeed$Character[walkspeed$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
walkspeed$Character[walkspeed$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
walkspeed$Character[walkspeed$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
walkspeed$Character[walkspeed$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
walkspeed$Character[walkspeed$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
walkspeed$Character[walkspeed$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
walkspeed$Character[walkspeed$Character == "Dedede"] <- "King Dedede"
walkspeed$Character[walkspeed$Character == "Metaknight"] <- "Meta Knight"
walkspeed$Character[walkspeed$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
walkspeed <- walkspeed %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(walkspeed, here::here("Project", "Data", "smash9_walkspeed.csv"))

# Load in Dash and Run Data
dashrun <- read.csv(here::here("Project", "Data", "smash10_dashrun.csv"))

# Fix Names
dashrun$Character[dashrun$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
dashrun$Character[dashrun$Character == "DK"] <- "Donkey Kong"
dashrun$Character[dashrun$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
dashrun$Character[dashrun$Character == "DK"] <- "Donkey Kong"
dashrun$Character[dashrun$Character == "P.T. [Charizard]"] <- "Charizard"
dashrun$Character[dashrun$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
dashrun$Character[dashrun$Character == "P.T. [Squirtle]"] <- "Squirtle"
dashrun$Character[dashrun$Character == "PAC-MAN"] <- "Pac-Man"
dashrun$Character[dashrun$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
dashrun$Character[dashrun$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
dashrun$Character[dashrun$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
dashrun$Character[dashrun$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
dashrun$Character[dashrun$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
dashrun$Character[dashrun$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
dashrun$Character[dashrun$Character == "Dedede"] <- "King Dedede"
dashrun$Character[dashrun$Character == "Metaknight"] <- "Meta Knight"
dashrun$Character[dashrun$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
dashrun <- dashrun %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(dashrun, here::here("Project", "Data", "smash10_dashrun.csv"))

# Load in Dash Turn Around Data
dashturn <- read.csv(here::here("Project", "Data", "smash11_dashturnaround.csv")) %>% 
  rename(
    Fast.Initial.Dash = Fast..Initial.Dash.,
    Normal.Buffer.Window = Normal..Buffer.Window.,
    Slow.Full.Dash = Slow..Full.Dash.
  )

# dashturn <- read.csv(here::here("Project", "Data", "smash11_dashturnaround.csv"))

# Fix Names
dashturn$Character[dashturn$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
dashturn$Character[dashturn$Character == "DK"] <- "Donkey Kong"
dashturn$Character[dashturn$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
dashturn$Character[dashturn$Character == "DK"] <- "Donkey Kong"
dashturn$Character[dashturn$Character == "P.T. [Charizard]"] <- "Charizard"
dashturn$Character[dashturn$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
dashturn$Character[dashturn$Character == "P.T. [Squirtle]"] <- "Squirtle"
dashturn$Character[dashturn$Character == "PAC-MAN"] <- "Pac-Man"
dashturn$Character[dashturn$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
dashturn$Character[dashturn$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
dashturn$Character[dashturn$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
dashturn$Character[dashturn$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
dashturn$Character[dashturn$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
dashturn$Character[dashturn$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
dashturn$Character[dashturn$Character == "Dedede"] <- "King Dedede"
dashturn$Character[dashturn$Character == "Metaknight"] <- "Meta Knight"
dashturn$Character[dashturn$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
dashturn <- dashturn %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(dashturn, here::here("Project", "Data", "smash11_dashturnaround.csv"))

# Load in Grab Data
grab <- read.csv(here::here("Project", "Data", "smash12_grab.csv"))

# Fix Names
grab$Character[grab$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
grab$Character[grab$Character == "DK"] <- "Donkey Kong"
grab$Character[grab$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
grab$Character[grab$Character == "DK"] <- "Donkey Kong"
grab$Character[grab$Character == "P.T. [Charizard]"] <- "Charizard"
grab$Character[grab$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
grab$Character[grab$Character == "P.T. [Squirtle]"] <- "Squirtle"
grab$Character[grab$Character == "PAC-MAN"] <- "Pac-Man"
grab$Character[grab$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
grab$Character[grab$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
grab$Character[grab$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
grab$Character[grab$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
grab$Character[grab$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
grab$Character[grab$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
grab$Character[grab$Character == "Dedede"] <- "King Dedede"
grab$Character[grab$Character == "Metaknight"] <- "Meta Knight"
grab$Character[grab$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
grab <- grab %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(grab, here::here("Project", "Data", "smash12_grab.csv"))

## Lots of changes for Ledge Stats
# Load in LEdge Stats Data
ledge <- read.csv(here::here("Project", "Data", "smash13_ledge.csv"))

# Fix Names
ledge$Character[ledge$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
ledge$Character[ledge$Character == "DK"] <- "Donkey Kong"
ledge$Character[ledge$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
ledge$Character[ledge$Character == "DK"] <- "Donkey Kong"
ledge$Character[ledge$Character == "P.T. [Charizard]"] <- "Charizard"
ledge$Character[ledge$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
ledge$Character[ledge$Character == "P.T. [Squirtle]"] <- "Squirtle"
ledge$Character[ledge$Character == "PAC-MAN"] <- "Pac-Man"
ledge$Character[ledge$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
ledge$Character[ledge$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
ledge$Character[ledge$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
ledge$Character[ledge$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
ledge$Character[ledge$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
ledge$Character[ledge$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
ledge$Character[ledge$Character == "Dedede"] <- "King Dedede"
ledge$Character[ledge$Character == "Metaknight"] <- "Meta Knight"
ledge$Character[ledge$Character =="Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
ledge <- ledge %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(ledge, here::here("Project", "Data", "smash13_ledge.csv"))

# Load in Out of Shield Data
outofshield <- read.csv(here::here("Project", "Data", "smash14_outofshield.csv"))

# Fix Names
outofshield$Character[outofshield$Character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
outofshield$Character[outofshield$Character == "DK"] <- "Donkey Kong"
outofshield$Character[outofshield$Character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
outofshield$Character[outofshield$Character == "DK"] <- "Donkey Kong"
outofshield$Character[outofshield$Character == "P.T. [Charizard]"] <- "Charizard"
outofshield$Character[outofshield$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
outofshield$Character[outofshield$Character == "P.T. [Squirtle]"] <- "Squirtle"
outofshield$Character[outofshield$Character == "PAC-MAN"] <- "Pac-Man"
outofshield$Character[outofshield$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
outofshield$Character[outofshield$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
outofshield$Character[outofshield$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
outofshield$Character[outofshield$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
outofshield$Character[outofshield$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
outofshield$Character[outofshield$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
outofshield$Character[outofshield$Character == "Dedede"] <- "King Dedede"
outofshield$Character[outofshield$Character == "Metaknight"] <- "Meta Knight"
outofshield$Character[outofshield$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Separate Echo Fighters
outofshield <- outofshield %>% 
  separate_rows(Character, sep = "/")

# Write data locally
write_csv(outofshield, here::here("Project", "Data", "smash14_outofshield.csv"))

```

Now to join all the datasets together. As well as clean some of the `Character` names since it is not consistent across `frame_data` and `tier_data`

```{r join data}

frame_data <- acceleration %>% 
  full_join(airspeed, by = "Character") %>% # Ice Climbers together
  full_join(fallspeed, by = "Character") %>%  # Sephiroth 
  full_join(gravity, by = "Character") %>% 
  full_join(jumpheight, by = "Character") %>% # Rosalina & Luma
  full_join(jumpduration, by = "Character") %>% 
  full_join(weight, by = "Character") %>% 
  full_join(landing, by = "Character") %>% 
  full_join(walkspeed, by = "Character") %>% 
  full_join(dashrun, by = "Character") %>% 
  full_join(dashturn, by = "Character") %>% 
  full_join(grab, by = "Character") %>% 
  full_join(ledge, by = "Character") %>% 
  full_join(outofshield, by = "Character")


unique(frame_data$Character[order(frame_data$Character)])

# Fix names
frame_data$Character[frame_data$Character == "DK"] <- "Donkey Kong"
frame_data$Character[frame_data$Character == "P.T. [Charizard]"] <- "Charizard"
frame_data$Character[frame_data$Character == "P.T. [Ivysaur]"] <- "Ivysaur"
frame_data$Character[frame_data$Character == "P.T. [Squirtle]"] <- "Squirtle"
frame_data$Character[frame_data$Character == "PAC-MAN"] <- "Pac-Man"
frame_data$Character[frame_data$Character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
frame_data$Character[frame_data$Character == "Pokemon Trainer (Charizard)"] <- "Charizard"
frame_data$Character[frame_data$Character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
frame_data$Character[frame_data$Character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
frame_data$Character[frame_data$Character == "Olimar (r/b/y)"] <- "Olimar (rby)"
frame_data$Character[frame_data$Character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
frame_data$Character[frame_data$Character == "Dedede"] <- "King Dedede"
frame_data$Character[frame_data$Character == "Metaknight"] <- "Meta Knight"
frame_data$Character[frame_data$Character == "Mii Swordfighter"] <- "Mii Sword Fighter"
frame_data$Character[frame_data$Character == "R.O.B."] <- "R.O.B"

frame_clean <- frame_data %>% 
  separate_rows(Character, sep = "/")

unique(frame_clean$Character[order(frame_clean$Character)])

write_csv(frame_clean, here::here("Project", "Data", "frame_data.csv"))

```

### Character Usage Data

I have found some data on the usage of the characters. Here is the [website](https://www.eventhubs.com/stats/ssbu/) I webscrapped the data from This shows a tier list and the player usage. I think even though this might not be a completely comprehensive list it does show how some characters are chosen over others. Since there are certain characters that are used more and I want to know if there is any underlying frame data that is causing these choices.

I also need to clean the `character` variable names so it matches the `frame_data`.

```{r webscraping usage}
# Webscraping
tables <- read_html("https://www.eventhubs.com/stats/ssbu/") %>%
  html_nodes("#tiers1")

# Practice table webscrape
tier_list <- tables %>%
  .[1] %>%
  html_table(fill = TRUE)

tier_data <- as.data.frame(tier_list) %>% 
  select(-X.) %>% 
  rename(
    character = Character.name,
    score = Score,
    total_players = Total.players,
    per_played = Percentage.played,
    string1 = X1S,
    string2 = X2S,
    string3 = X3S
  ) %>% 
  separate_rows(character, sep = " / ")

# Look at all names in alphabetical order to see what ones are spelled differently
unique(tier_data$character[order(tier_data$character)])

# Fix names
tier_data$character[tier_data$character == "Banjo-Kazooie"] <- "Banjo & Kazooie"
tier_data$character[tier_data$character == "DK"] <- "Donkey Kong"
tier_data$character[tier_data$character == "Mr. Game and Watch"] <- "Mr. Game & Watch"
tier_data$character[tier_data$character == "P.T. [Charizard]"] <- "Charizard"
tier_data$character[tier_data$character == "P.T. [Ivysaur]"] <- "Ivysaur"
tier_data$character[tier_data$character == "P.T. [Squirtle]"] <- "Squirtle"
tier_data$character[tier_data$character == "PAC-MAN"] <- "Pac-Man"
tier_data$character[tier_data$character == "Pokemon Trainer (Ivysaur)"] <- "Ivysaur"
tier_data$character[tier_data$character == "Pokemon Trainer (Charizard)"] <- "Charizard"
tier_data$character[tier_data$character == "Pokemon Trainer (Squirtle)"] <- "Squirtle"
tier_data$character[tier_data$character == "Olimar (r/b/y +2)"] <- "Olimar (rby +2)"
tier_data$character[tier_data$character == "Olimar (r/b/y)"] <- "Olimar (rby)"
tier_data$character[tier_data$character == "Olimar (r/b/y +1)"] <- "Olimar (rby +1)"
tier_data$character[tier_data$character == "Dedede"] <- "King Dedede"
tier_data$character[tier_data$character == "Metaknight"] <- "Meta Knight"
tier_data$character[tier_data$character == "Mii Swordfighter"] <- "Mii Sword Fighter"

# Double check names
unique(tier_data$character[order(tier_data$character)])
  
write_csv(tier_data, here::here("Project", "Data", "tier_data.csv"))

```


## Merging Data

Now that I have `frame_data` and `tier_data` I can merge them based on the character name and then the `per_played` will be my outcome variable with the frame data being the explanatory variables. I am going to use `full_join` again. I need to fix the issue with the echo fighters being combined on the `tier_data` but separate in the `frame_data`.

```{r join frame and usage}
# Read in Data
frame_data <- read_csv(here::here("Project", "Data", "frame_data.csv"))

tier_data <- read_csv(here::here("Project", "Data", "tier_data.csv"))

# Join Datasets
smash_data <- tier_data %>% 
  right_join(frame_data, by = c("character" = "Character"))

unique(smash_data$character[order(smash_data$character)])

# Write Data Locally
write_csv(smash_data, here::here("Project", "Data", "smash_data.csv"))
```

Data cleaning is pretty much done. There are a few weird instances of character name difference that I don't care about at this moment. Also the `gravity` dataset might be wrong, but I don't think I will use any of those variables anyways.

### Failed Data Gathering 

Originally I wanted to access match results to use as my outcome variable. However I was unable to do this and here is some of the code of my attempts to connect to a database I found on github.

Now that I have the frame data collected. I can get the Match data which will tell me the wins and losses of each character. My goal is to create a model that will determine if a character wins or losses. Maybe a win rate or discrete explanatory variable we will see.

```{r connect to database}
# library(DBI)
# library(odbc)
# # Create an ephemeral in-memory RSQLite database
# con <- dbConnect(RSQLite::SQLite(), dbname = "ultimate_player_database.db")
# 
# dbListTables(con)
# 
# con <- dbConnect(RSQLite::SQLite(), dbname = "ultimate_player_database.db")
# 
# con <- dbConnect(
#   RPostgreSQL::PostgreSQL(), "ultimate_player_database.db",
# )
# 
# dbListTables(con)
# 
# ## Connect to an existing database
# test <- dbConnect(RSQLite::SQLite(), 'https://github.com/smashdata/ThePlayerDatabase/releases/download/v2021.02.05/ultimate_player_database.db')
# 
# dbListTables(test)
# 
# file.exists("https://github.com/smashdata/ThePlayerDatabase/releases/download/v2021.02.05/ultimate_player_database.db")
# 
# file.info("https://github.com/smashdata/ThePlayerDatabase/releases/download/v2021.02.05/ultimate_player_database.db")
# 
# con <- dbConnect(odbc(), "ultimate_player_database.db")

```

```{r}
# library(RSQLite)
# temp <- tempfile()
# # download.file("https://github.com/smashdata/ThePlayerDatabase/releases/download/v2021.02.05/ultimate_player_database.db", temp)
# db <- dbConnect(RSQLite::SQLite(), dbname=temp)
# dbListTables(db)

```

```{r}
# # Packages
# library(tidyverse)
# library(httr)
# library(jsonlite)
#
# # # Save API Key to Private Folder
# # saveRDS(smash_key, here::here("Private", "smash_key.rds"))
# 
# # Read in API Key
# api_key <- readRDS(here::here("Private", "smash_key.rds"))
# 
# # Url to API
# url <- "https://api.smash.gg/gql/alpha"
# 
# resp <- httr::GET(url,
#           add_headers("Bearer" = api_key))
# 
# httr::GET(url,
#           add_headers("Bearer" = api_key))
# 
# # Check for any errors
# http_error(resp) 



```

