---
title: "Assignment 9"
author: "Zack Wixom"
date: "3/27/2021"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/Week 11/"
)
```

```{r packages, message = FALSE}
# Packages
library(tidyverse)
library(rethinking)
library(dagitty)
library(ellipse)

# Set Seed
set.seed(42)

# Load Data
data("bangladesh")
d <- bangladesh

```

### Question 1

First to make the varying slopes model, with centered parameterization.

```{r m1.1 code}
# Standardize Vars
data <- tibble(
  contra = d$use.contraception,
  district_id = as.integer(as.factor(d$district)),
  urban = d$urban
)

# Fit Model
m1.1 <- ulam(
  alist(
    contra ~ bernoulli(p),
    logit(p) <- a[district_id] + b[district_id]*urban,
    c(a, b)[district_id] ~ multi_normal(c(abar, bbar), Rho, Sigma),
    abar ~ normal(0, 1),
    bbar ~ normal(0, 0.5),
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1)
  ),
  data = data,
  chains = 4,
  cores = 4,
  cmdstan = TRUE
)

```

```{r m1.1 plot 1}
plot(precis(m1.1))
```

This is telling us that urban areas are using more contraceptives. But we will look at distribution of varying effects.

```{r m1.1 plot 2}

plot(precis(m1.1, depth = 3, pars = c("Rho", "Sigma")))

```

**Can't remember what differeence of looking at Rho does, Check in Book**

Now onto plotting the intercpets and slopes with the ellipse plot 

```{r m1.1 ellipse}
post <- extract.samples(m1.1)

a <- apply(post$a, 2, mean)
b <- apply(post$b, 2, mean)

plot(a, b, xlab = "a intercept", ylab = "b urban slope")
abline(h = 0, lty = 2)
abline(v = 0, lty = 2)

R <- apply(post$Rho, 2:3, mean)
s <- apply(post$Sigma, 2, mean)
S <- diag(s) %*% R %*% diag(s)
ll <- c(0.5, 0.67, 0.89, 0.97)
for(l in ll){
  el <- ellipse(S, contre = c(mean(post$abar), mean(post$bbar)), level = l)
  lines(el, col = "black", lwd = 0.5)
}



```

So we can see a negative correlation with districts with higher use outside urban areas have smaller slopes. 

Overall this can mean that perhaps urban areas are similar to all districts but once you go to rural areas there is variation, But to check we will plot urban areas against rural areas.

```{r m1.1 outcome scale}
u0 <- inv_logit(a)
u1 <- inv_logit(a + b)

plot(u0, u1, xlim = c(0,1), ylim = c(0,1), xlab = "rural", ylab = "urban")
abline(h = 0.5, lty = 2)
abline(v = 0.5, lty = 2)

```

This shows how the spread is more tight in the urban areas while the rural as a large spread and more variety.

### Question 2

First to draw my DAG with `A` as age, `K` as number of children, and `C` as contraception.

```{r dag}
# Create DAG
dag1 <- dagitty("dag{A -> K; K -> C; A -> C}")

# DAG coordinates
coordinates(dag1) <- list(x = c(A = 0, C = 1, K = 2), y = c(A = 0, C = 1, K = 0))

# Draw DAG
drawdag(dag1)

```

Now onto the modeling

```{r m2.1, message = FALSE}
# Standardize Vars
data <- tibble(
  contra = d$use.contraception,
  district_id = as.integer(as.factor(d$district)),
  urban = d$urban,
  children = standardize(d$living.children),
  age = standardize(d$age.centered)
)

# Fit Model
m2.1 <- ulam(
  alist(
    contra ~ bernoulli(p),
    logit(p) <- a[district_id] + b[district_id]*urban + bA*age,
    c(a, b)[district_id] ~ multi_normal(c(abar, bbar), Rho, Sigma),
    abar ~ normal(0, 1),
    bbar ~ normal(0, 0.5),
    bA ~ normal(0, 0.5),
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1)
  ),
  data = data,
  chains = 4,
  cores = 4,
  cmdstan = TRUE
)

```

```{r m2.1 plot}
plot(precis(m2.1))
```

So it looks like age doesn't give us a really big impact, it is slightly positive though.

So let's add in `K` as number of children.

```{r m2.2}
# Fit Model
m2.2 <- ulam(
  alist(
    contra ~ bernoulli(p),
    logit(p) <- a[district_id] + b[district_id]*urban + bA*age + bK*children,
    c(a, b)[district_id] ~ multi_normal(c(abar, bbar), Rho, Sigma),
    abar ~ normal(0, 1),
    bbar ~ normal(0, 0.5),
    c(bA,bK) ~ normal(0, 0.5),
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1)
  ),
  data = data,
  chains = 4,
  cores = 4,
  cmdstan = TRUE
)

```

```{r m2.2 plot}
plot(precis(m2.2))
```

Now we can see there is a negative effect of age once number of children is conditioned and that has a positive effect on it. I guess that means that people use more contraceptives once they have more children and the older they are the less likely the are to use contraceptives. I guess the older the parents the more likely they will have more children over their life time. 

### Question 3

No we will add the ordered categorical variables.

```{r m3.1, message = FALSE}
# Standardize Vars
data <- list(
  contra = d$use.contraception,
  district_id = as.integer(as.factor(d$district)),
  urban = d$urban,
  children = d$living.children,
  age = standardize(d$age.centered)
)

data$alpha <- rep(2,3)

# Fit model
m3.1 <- ulam(
  alist(
    contra ~ bernoulli(p),
    logit(p) <- a[district_id] + b[district_id]*urban + bA*age + bK*sum(delta_shell[1:children]),
    c(a, b)[district_id] ~ multi_normal(c(abar, bbar), Rho, Sigma),
    abar ~ normal(0, 1),
    bbar ~ normal(0, 0.5),
    c(bA,bK) ~ normal(0, 0.5),
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1),
    vector[4]: delta_shell <<- append_row(0, delta),
    simplex[3]: delta ~ dirichlet(alpha)
  ),
  data = data,
  chains = 4,
  cores = 4,
  cmdstan = TRUE
)


```

```{r m3.1 plot}

plot(precis(m3.1))

```

So we can see that it is similiar to `m2.2`. Age slightly reduces use of contraceptives while the number of kids increases use.

let's plot with `pars = "delta"`.

```{r m3.1 plot delta}

plot(precis(m3.1, 3, pars = "delta"))


```

Looks like the change from having 1 to 2 kids greatly increases the use of contraceptives. 
