---
title: "Week 2 Notes"
author: "Zack Wixom"
output: github_document
---

## Golem of Prague

Computers are like _Golems_. They do what we tell them but can end up destroying things because it doesn't know our intent. 

## Installing STAN

```{r install stan}
# # we recommend running this is a fresh R session or restarting your current session
# install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```

Now that Stan is installed we can call `cmdstanr` package to use Stan, as well as `posterior` and `bayesplot`

```{r}
# Packages
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
color_scheme_set("brightblue")

# Use check_cmdstan_toolchain() to see if cmdstanr is working

# then install cmdstanr with
install_cmdstan(cores = 2)
```

## Statistical Rethinking package

Now to install the package that goes along with the book we are using

```{r}
install.packages(c("coda", "mvtnorm", "devtools", "loo"))
devtools::install_github("rmcelreath/rethinking")
```


# What is Bayesian Data Analysis

Using probability as a language to describe uncertainty

- Extends ordinary binary logic (True/False) to continuous plausibilities

Difficult to do so we use MCMC (Markov chain Monte Carlo)

Used to be controversial in the stats world

Into one statement:

__Count all the ways data can happen, according to your assumptions. The assumptions with more ways that are consistent with data are more plausible.__

## Multi-level Models

__Models within models__

- Models with multiple levels of uncertainty
- Replace parameters with models

__Common uses__

- Repeat & Imbalance sampling
- Study variation
- Avoid averaging
- Phlyogenetics, factor and path analysis, networks and spatial models

Natural Bayesian strategy






