---
title: "Assignment 5"
subtitle: "MCMC"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/"
)
```

```{r packages, message=FALSE}
# Packages
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(rethinking)

# Load Data
data("Wines2012")

d <- Wines2012

```

## Question 1

```{r q1 standardize and index vars}
# Standardize  and Index Variables
wines <- list(
    score = standardize(d$score),
    judge_id = as.integer(d$judge),
    wine_id = as.integer(d$wine)
)
```

Not sure if I need to also do the log and std of the outcome variable or not. I tried and it wasn't working so I will keep with just the original standardized `score` variable.

```{r q1 linear model, message = FALSE}
# Fit model
m1.1 <- quap(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a[judge_id] + b[wine_id],
    a[judge_id] ~ dnorm(0, 0.5),
    b[wine_id] ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ), data = wines
)

```

```{r q1.1 precis}
plot(precis(m1.1, 2))
```

```{r q1 mcmc, message = FALSE}
# Fit Model
m1.2 <- ulam(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a[judge_id] + b[wine_id],
    a[judge_id] ~ dnorm(0, 0.5),
    b[wine_id] ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data=wines, 
  chains=4, 
  cores = 4
)

```

```{r, q1.2 plot}
plot(precis(m1.2, depth = 2))
```

__How do you interpret the variation among individual judges and individual wines?__

I am assuming that different judges have different preferences for certain wines.

__Do you notice any patterns, just by plotting the differences?__

I wasn't able to get to get the plot to work, perhaps my priors are off as well. But I do see there are a lot of differences between the judges and not the wines, maybe suggesting that there isn't much difference between them.

__Which judges gave the highest/lowest ratings?__

```{r q1 judge summary}
d %>% 
  group_by(judge) %>% 
  summarize(
    id = max(as.integer(judge))
  )
```

_Lowest_:  Robert Hodgson, Jean-M Cardebat, and Tyler Colman.

_Highest_: John Foy and Linda Murphy.

__Which wines were rated worst/best on average?__

```{r q1 wine summary}
d %>% 
  group_by(wine) %>% 
  summarize(
    id = max(as.integer(wine))
  )
```

_Highest wines:_ B2 and J2

_Lowest wines:_ I2 and C2

## Question 2

```{r q2 standardize}
# Standardize  and Index Variables
wines2 <- list(
    score = standardize(d$score),
    wine_id = d$wine.amer + 1L,
    judge_id = d$judge.amer + 1L,
    flight_id = as.integer(d$flight)
)

```

```{r q2 mcmc, message = FALSE}
# Fit Model
m1.2 <- ulam(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- w[wine_id] + j[judge_id] + f[flight_id],
    w[wine_id] ~ dnorm(0, 0.5),
    j[wine_id] ~ dnorm(0, 0.5),
    f[wine_id] ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ), 
  data = wines2, 
  chains = 4, 
  cores = 4
)

```

```{r q2 plot}
plot(precis(m1.2, depth = 2))
```

__What do you conclude about the differences among the wines and judges?__

It looks like there aren't differences between the red and white wine scoring but there is a noticeable difference between where the judge and wine are from. American judges will score higher than French. However, the French wines score higher than American wines.

## Question 3

For this question I am doing a different approach. Not sure which one is the best in general but I think it is better for an indicator approach to the predictor variables, than mbefore I was using index variables.

```{r q3 standardize}
# Standardize and M<ake Indicator Variables
wines3 <- list(
    score = standardize(d$score),
    wine = d$wine.amer,
    judge = d$judge.amer,
    flight = ifelse(d$flight == "red", 1, 0)
)
```

For the interactions I am doing three different ones: _wine & judge, wine & flight,_ and _judge & flight._

I saw something that suggested to do interaction prior tighter so that's why I chose these priors.

```{r q3 mcmc, message = FALSE}
# Fit Model
m1.3 <- ulam(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a + 
      bW*wine + 
      bJ*judge + 
      bF*flight + 
      bWJ*wine*judge + 
      bWF*wine*flight + 
      bJF*judge*flight,
    a ~ dnorm(0, 0.2),
    bW ~ dnorm(0, 0.5),
    bJ ~ dnorm(0, 0.5),
    bF ~ dnorm(0, 0.5),
    bWJ ~ dnorm(0, 0.25),
    bWF ~ dnorm(0, 0.25),
    bJF ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), 
  data = wines3, 
  chains = 4, 
  cores = 4
)

```

```{r q3 plot}
plot(precis(m1.3))
```

Now to use `link()` to interpret the models predictions

```{r q3 link}
# Predictive data
pred_data <- tibble(
  wine = rep(0:1, times = 4),
  judge = rep(0:1, each = 4),
  flight = rep(c(0,0,1,1), times = 2)
)

# Link
mu <- link(m1.3, data = pred_data)

# Row names for plot easier interpretation (found a resource to add these feature made it easier)
row_labels <- paste(ifelse(pred_data$wine == 1, "A", "F"),
                    ifelse(pred_data$judge == 1, "A", "F"),
                    ifelse(pred_data$flight == 1, "R", "W"), 
                    sep = "")

# Plot precis()
plot(precis(list(mu = mu), 2), labels = row_labels)
```

__What do you conclude about the features and the scores? Can you relate the results of your model(s) to the individual judge and wine inferences from Problem 1?__

Now with labels I can read this better and see that American Red Wine that scores low from French judge. While French Red Wine scores high from American Judge.

In general this is similar to my inferences from problem one. It looks like French judges score low on average and american score higher. French wines score higher compared to American ones, but there isn't much difference between Red and White wines that I can realy tell here.



