---
title: "Assignment 7"
author: "Zack Wixom"
output: github_document
---

```{r opts, echo = FALSE}
# puts any figures made into this folder
knitr::opts_chunk$set(
  fig.path = "../Figures/Week 9/"
)
```

```{r packages, message = FALSE}
# Packages
library(rethinking)
library(tidyverse)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(dagitty)

data(Trolley)

d <- Trolley

# Recode Education
edu_levels <- c(6, 1, 8, 4, 7, 2, 5, 3)
d$edu_new <- edu_levels[d$edu]

# Standardize Vars
idx <- 1:nrow(d)
dat <- list(
  y = d$response[idx],
  A = d$action[idx],
  I = d$intention[idx],
  C = d$contact[idx],
  E = as.integer(d$edu_new[idx]),
  edu_norm = normalize(d$edu_new[idx]),
  age = standardize(d$age[idx]),
  alpha = rep(2, 7)
)
```

## Question 1

Here is my DAG:

```{r DAG 1}
dag1 <- dagitty( "dag{
    E -> R
    A -> R
    A -> E
}")

# Coordinates for upside down triangle
coordinates(dag1) <- list(x = c(E = 0, R = 1, A = 2), y = c(E = 0, R = 1, A = 0))

# Draw DAG
drawdag(dag1)

```

Now for the model. I already pre-processed the data above question 1.

```{r model 1, message = FALSE}
m1 <- ulam(
  alist(
    y ~ ordered_logistic(phi, cutpoints),
    phi <- bE*sum(delta_shell[1:E]) + bA*A + bC*C + BI*I + bAge*age,
    # Add interactions
    BI <- bI + bIA*A + bIC*C,
    c(bA, bI, bC, bIA, bIC, bE, bAge) ~ normal(0, 0.5),
    cutpoints ~ normal(0, 1.5),
    vector[8]: delta_shell <<- append_row(0, delta),
    simplex[7]: delta ~ dirichlet(alpha)
  ),
  data = dat,
  chains = 4, 
  cores = 4,
  cmdstan = TRUE
)


```

```{r model 1 precis}
precis(m1)
```

Looks like Education has the biggest positive impact on choice. Age has a little bit of a negative cause. Overall it looks like people with more education can make better decisions.

## Question 2

First to draw my DAG but including Gender.

```{r DAG 2}
dag2 <- dagitty( "dag{
    E -> R
    A -> R
    A -> E
    G -> R
    G -> E
}")

# Coordinates for upside down triangle
coordinates(dag2) <- list(
  x = c(E = 0.5, R = 1.5, A = 2, G = 0), 
  y = c(E = 0, R = 1, A = 0, G = 1)
)

# Draw DAG
drawdag(dag2)

```

We need to condition on both A and G to make sure there is not a confound from E.

Now onto the model.

```{r model 2, message = FALSE}
# Make indicator for gender
dat$female <- ifelse(d$male == 1, 0L, 1L)

# Model
m2 <- ulam(
  alist(
    y ~ ordered_logisitic(phi, cutpoints),
    phi <- bE*sum(delta_shell[1:E]) + bA*A + bC*C + BI*I + bAge*age + bF*female,
    BI <- bI + bIA*A + bIC*C,
    c(bA, bI, bC, bIA, bIC, bE, bAge, bF) ~ normal(0. 0.5),
    cutpoints ~ normal(0, 1.5),
    vector[8]: delta_shell <<- append_row(0, delta),
    simplex[7]: delta ~ dirichlet(alpha)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  cmdstan = TRUE
)

```

```{r}
precis(m2)
```


